
\documentclass[11pt]{exam} % https://www.ctan.org/pkg/exam?lang=en

\usepackage[lmargin=1.in,rmargin=1.in,tmargin=1.in,bmargin=1in]{geometry}
\usepackage{setspace}
\usepackage[pdftex]{graphicx}
\usepackage{titling}
\usepackage[
	pdfauthor={Brian Weinstein},
	pdftitle={Homework 1},
	bookmarks=true,
	colorlinks=true,
	linkcolor=blue,
	urlcolor=blue,
	citecolor=blue,
	pdftex,
	linktocpage=true
	]{hyperref}
\usepackage[textsize=tiny]{todonotes}
\usepackage{float}
\setlength\parindent{0pt}
\usepackage{lipsum}
\usepackage{amsmath}


\qformat{\textbf{Problem \thequestion: \thequestiontitle}\quad \hfill}


\pagestyle{headandfoot}
\runningheadrule
\firstpageheader{}{}{}
\runningheader{\theauthor}{\thetitle}{\thedate}
\firstpagefooter{}{\thepage}{}
\runningfooter{}{\thepage}{}


\usepackage{xcolor}
\usepackage{adjustbox}
\usepackage{verbatim}
\definecolor{shadecolor}{rgb}{.9, .9, .9}

\newenvironment{code}%
   {\par\noindent\adjustbox{margin=1ex,bgcolor=shadecolor,margin=0ex \medskipamount}\bgroup\minipage\linewidth\verbatim}%
   {\endverbatim\endminipage\egroup}

\newenvironment{codeSmall}%
   {\par\noindent\adjustbox{margin=1ex,bgcolor=shadecolor,margin=0ex \medskipamount}\bgroup\minipage\linewidth\verbatim\footnotesize}%
   {\endverbatim\endminipage\egroup}

\newcommand{\ramsey}{\href{http://www.statisticalsleuth.com/}{Ramsey }}



\begin{document}


\title{STAT W4201 001, Homework 3}
\author{Brian Weinstein (bmw2148)}
\date{Feb 17, 2016}
\maketitle

Code is attached here and also posted at \href{https://github.com/BrianWeinstein/advanced-data-analysis}{https://github.com/BrianWeinstein/advanced-data-analysis}. Where relevant, code snippets and output are are included in-line.

\begin{questions}


\titledquestion{\ramsey 4.30} % Problem 1


\titledquestion{\ramsey 4.32} % Problem 2



\titledquestion{\ramsey 5.19} % Problem 3

\begin{parts}

\part The pooled estimate of the variance $s_p^2$ is given by

\begin{equation*}
\begin{split}
s_p^2 &= \frac{\sum_{i=1}^{I} (n_i -1 ) s_i^2}{\sum_{i=1}^{I} (n_i -1 )} \\
&= \frac{   (127-1)(0.4979)^2  +  (44-1)(0.4235)^2  +  (24-1)(0.3955)^2  +  \cdots  }{(127-1) +  (44-1) +  (24-1)  +  \cdots} \\
& \quad \quad \quad \frac{\cdots + (41-1)(0.3183)^2 + (18-1)(0.3111)^2  +  (16-1)(0.4649)^2 + \cdots}{\cdots + (41-1) + (18-1)  +  (16-1) + \cdots } \\
& \quad \quad \quad \frac{\cdots  +  (11-1)(0.2963)^2  +  (7-1)(0.3242)^2  +  (6-1)(0.5842)^2  }{\cdots +  (11-1) +  (7-1)  +  (6-1)} \\
&= 0.1919.
\end{split}
\end{equation*}

\part To construct an ANOVA table to test for species differences, we first compute $SS_W$, the sum of squared residuals of the ``within'' group.

\begin{align*}
SS_W &= \sum_{i=1}^9 \left[ \sum_{j=1}^{n_i} \left( Y_{ij} - \overline{Y_i} \right)^2  \right] = \sum_{i=1}^9 \left[ (n_i - 1) \text{SE}(Y_i)  \right] \\
&= (127 - 1)(0.4979) + (44 - 1)(0.4235) + \cdots + (6 - 1)(0.5842) \\
&= 122.8658.
\end{align*}

Given the sample standard deviation of all 294 observations as one group is $\text{SD}_{Total} = 0.4962$, the total sum of squared residuals $SS_{Total}$ is

\todo{Should SS total be -9 (-I) instead of -1 ?}
\begin{align*}
SS_{Total} &= \sum_{i=1}^9 \left[ \sum_{j=1}^{n_i} \left( Y_{ij} - \overline{Y} \right)^2  \right] = (n_1 + n_2 + \cdots + n_9 - 1) (\text{SD}_{Total}) \\
&= (127 + 44 + \cdots + 6 - 1)(0.4962) =  (293)(0.4962)\\
&= 145.3866.
\end{align*}

Therefore, the sum of squared residuals of the ``between'' group, $SS_B$ is:
$$SS_B = SS_{Total} - SS_W = 145.3866 - 122.8658 = 22.5208.$$

The Total, Within, and Between degrees of freedom are $\text{df}_{Total} = (n-1) = 293$, $\text{df}_W = (n-I)=(294-9) = 285$, and $\text{df}_B = \text{df}_{Total} - \text{df}_W  = 8$, respectively. \\

The Mean Square for the Within and Between groups is the [(Sum of Squares) / df], and the F-statistic is defined as the [(Between Mean Square) / (Within Mean Square)]. \\

Thus the ANOVA table to test for species differences is:

\begin{center}
  \begin{tabular}{ l c c c c c}
    Source of Variation & Sum of Squares & d.f. & Mean Square & F-Statistic & p-Value \\ \hline \hline
    Between Groups & 22.5208 & 8 & 2.8151 & 6.5300 & $8.0551 \times 10^{-8}$ \\
    Within Groups & 122.8658 & 285 & 0.4311 & & \\ \hline
    Total & 145.3866 & 293 & & &
  \end{tabular}
\end{center}

\quad \linebreak
Where the p-value for the F-Statistic is given by the CDF of an F distribution with 8 and 285 degrees of freedom.
\begin{codeSmall}
> # p-value of F-statistic
> pf(q=6.5300, df1=8, df2=285, lower.tail=FALSE)
[1] 8.055137e-08
\end{codeSmall}



\part The ANOVA method for calculating $SS_B$ yields the same answer as the formula:

$$ SS_B = \sum_{i=1}^I \left( n_i \overline{Y_i}^2 \right) - n\overline{Y}^2 = \sum_{i=1}^9 \left( n_i \overline{Y_i}^2 \right) - n\overline{Y}^2 .$$

The overall mean $\overline{Y}=7.4746$ is computed as a weighted average of the group means and $n=294$, thus \todo{17.45402 != 22.5208 : maybe my SS total calc is incorrect?}
\begin{align*}
SS_B &= \sum_{i=1}^9 \left( n_i \overline{Y_i}^2 \right) - n\overline{Y}^2 \\
&= (127)(7.347)^2 + (44)(7.368)^2 + (24)(7.418)^2 + (41)(7.487)^2 + (18)(7.563)^2 + \cdots \\
& \quad \quad \quad \cdots + (16)(7.568)^2 + (11)(8.214)^2 + (7)(8.272)^2 + (6)(8.297)^2 - (294)(7.4746)^2 \\
&= 17.45402
\end{align*}


\part If the first 6 species have one common mean and the last 3 have another common mean, we find that \todo{finish 3d}




\end{parts}







\titledquestion{} % Problem 4
\textit{Consider the Bumpus’s data in Chapter 2, compute the power of the two-sided two sample t-test of size 0.05 (i.e., reject the null hypothesis if the absolute value the t-statistic is greater than or equal to 2), under the alternative that $\mu_x - \mu_y = \overline{x} - \overline{y} = 0.01$ and $\sigma = s_p = 0.0214$.}


\titledquestion{} % Problem 5
\textit{Show that the two-sided two sample t-test is equivalent to the anova F-test, if the number of groups is two.}

For $I=2$ groups, the F-statistic is given by
$$ \text{F-statistic} =  \frac{SS_B / \left[ (n-1) - (n-I) \right]}{SS_W / (n-I)}, $$
where $n_1$ and $n_2$ are the sizes of samples 1 and 2, respectively, $n=n_1+n_2$ is the total sample size, $SS_B$ is the ``between groups'' sum of squared residuals, and $SS_W$ is the ``within groups'' sum of squared residuals.

Simplifying, we find
$$\text{F-statistic} =  \frac{SS_B / (I-1) }{SS_W / (n-I)} = \frac{SS_B / (2-1) }{SS_W / (n-2)} = \frac{SS_B / 1 }{SS_W / (n-2)}.$$

If the observations from group 1 are $\sim \text{N}(\mu_1, \sigma^2)$ and the observations from group 2 are $\sim \text{N}(\mu_2, \sigma^2)$, we know that
$$\text{F-statistic} \sim \text{F}_{1, n-2} \text{ , which is equivalent to }t^2_{n-2}.$$
i.e., an F distribution with a numerator degrees of freedom of 1 and a denominator degrees of freedom of $n-2$ is equivalent to the square of a t distribution with $n-2$ degrees of freedom. \todo{Finish problem 5}



\titledquestion{} % Problem 6
\textit{Consider $X_1,\ldots,X_{10}$ are i.i.d. N($0,\sigma^2$), $Y_1,\ldots,Y_{10}$ are i.i.d. N($\mu,\sigma^2$) and hypothesis testing:}
\begin{align*}
H_0 &: \mu = 0\\
H_A &: \mu \neq 0.
\end{align*}
\textit{Compute the power of a two sided two sample t-test of size 0.05 when $\sigma^2 = 1$ and $\mu =$ 0.1, 0.5, 1, and 2. Plot the power as a function of $\mu$. Then, increase the sample size in each group to 20 and draw the power function in the same plot as that of the sample size 10.}



\titledquestion{} % Problem 7
\textit{Under the setting of the previous problem, show that, under the null hypothesis, the p-value follows the uniform distribution on the interval [0, 1] and perform simulations to confirm it.}




\end{questions}

\listoftodos

\end{document}